\chapter{Events}


The game handles actions by giving each thing a few events that get
triggered every few seconds for it.

\section{Thing Events}
 Most things get some of the
following events:

\begin{itemize}



\item EVENT\_THING: This happens every 7-11 seconds depending on how
the game is currently coded at the moment. This event deals with
regenerating health and also for actions like society
actions. Generally only mobs get this event, but other things that
cast spells or have some special activity may get it also.

\item EVENT\_COMBAT: This happens every 1.5-2 seconds for each thing
that is in combat. This also calls the hunting code since these two
things are mutually exclusive (in any given pulse) so it saves some
resources over having a separate hunting event. Only mobs that have
been in combat or have had to hunt something get this event.

\item EVENT\_THING\_HOUR: This happens every 60-80 seconds within the game
and it determines when tracks fade, and affects fade, and when things
that are timed delete themselves. Every thing gets this.

\item EVENT\_CAST: This is used by things that cast so that their
casting is independent of other activities. They must have a VAL\_CAST
on them for this to be set.
 
\item EVENT\_FAST: Certain things happen only around pc's such as
mobs doing socials, and mobs attacking faster. This event gets called
every couple of seconds for each player.


\end{itemize}

There is also an event called EVENT\_TRACK that players get to
use. This sets up a delay that lets them have a delay between entering
a room and finding tracks.

EVENT\_COMMAND is used to delay commands for players and mobs. One
example of players using this is the cast command. One example of
this for mobs is the wake command used when sleeping mobs wake up
after hearing a yell. They are staggered so that it looks like they
all wake up at different times.

An interesting idea may be to make haste/slow a number that gets
moved up and down around 100 (= normal) representing how hasted or
slow a thing is, and then reschedule these events faster or slower
based on that. I may do that at some point, but it would be hard to
balance, so it will be left for later.

\section{Global Events}
There are also several global events that happen at certain
intervals:

\begin{itemize}


\item EVENT\_AUCTION: This updates any auctions approximately every 15
seconds.

\item EVENT\_HOUR: This does updates like the hourly weather update
and some saving.

\item EVENT\_SOCIETY: A small percentage of societies get updated
every second to spread out the load. This calls that update function.

\item EVENT\_ALIGN: This updates the alignments to spread out the load
required to update their AI and data.


\end{itemize}

\section{Special Events}

There are also a couple of other special events.

EVENT\_DAMAGE: This event is used to deal out damage over time instead
of all at once. Spells have some ability to cause damage over time
now.

EVENT\_FREE\_THING: This is an event used to free things about a minute
after they are removed from the game. The reason for this staggering
of the destruction of things is to save on CPU. 

First, when a mob
dies, it leaves a lot of tracks and it's too expensive to search every
room for tracks of that mob, so every hour, rooms delete tracks that
belong to things that aren't in anything (such as deleted mobs). Since
these updates are staggered over the minute or so each hour of game
time takes up, it spreads out this cost, and doesn't duplicate as much
work.

Second, the code that saves the world snapshot works on its own
thread. This was done because I don't see any good cheap way of saving
a snapshot of the world other than just farming it off to its own
thread. If I left it all in the main thread, there would be a
noticeable lag, or the saving would have to be spread out so
there would be more problems with contamination. Otherwise, I could
use a DB, but that would be too expensive for the resources I have at
hand (and this isn't written that way). So, I save things using a
separate thread, and I do some checking in move\_thing, so that if a
thing that isn't saved tries to move into something that is saved
(except when pc's are around), the move fails. This happens about 5-10
times when I try to save 100000 items (with the game really sped up),
so I think that captures most of the errors that are likely to occur.
The reason things aren't deleted right away is that if they are 
removed from the game in the middle of them being saved by the other
thread, pointers may go bad in the middle of that, and strings may
get freed as they are being written to a file. That's not good, so
I just try to remove the item partially, then remove it fully about a
minute later.

Like I said, it isn't a perfect solution, but it's close enough to
what I want and it's cheap, so I can do more of the simulation that I
like (since I don't really care if the simulation is perfect...since
it's a numbers game to me).




