\chapter{History Generation}

\section{Introduction}

There are several parts to generating the history of the world. This stuff has
to happen after the rest of the world is generated because it requires a lot
of data from those sections.

\section{Deities}

Each alignment gets a certain number of deities created. Basically, I made a
list of all kinds of "spheres of influence" that I thought deities might have
and split them into two lists: good and bad. The lists are in the historygen
area (currently 109500) and they're called godly\_spheres\_good and
godly\_spheres\_bad. Most deities are good, but some are bad. The game just
picks 2-4 different spheres for each deity. It actually looks pretty good. If
you want to try this out several times, you can type historygen setup and it
will make the gods and the races.

{\bf{Deities can be summoned by using the "divine" command and when they
    appear, they're level 500 and kill all enemies in sight. This divine skill
    requires a highlevel healer/mystic and it requires 5000 warpoints, so it
    should be used rarely. However, deities load with a FULL set of level 500
    equipment, so eventually players will kill the deities and get this eq and
    be really cheater. But I guess everyone needs a hobby. :)}}

\section{Organizations}

Each alignment has several societies of middling power that can help
them. They are named in the form "The foo of bars" where foo is something like
organization or company, and bar is some profession name like fighter or wizard.
These things are fairly powerful and can be powerful allies, but they can be
destroyed and turned to other sides.

\section{Ancient Races}

These are VERY powerful societies and each alignment has one. For now, they're
in the game to start, but they may end up being taken out and requiring
players to quest to get them on their sides.

\section{History Generation}

Ok, so most stories in online games suck. Let's face it, the one story that
gets used over and over again in online and offline RPG's is the "Heroes'
Quest" which succinctly can be described as:

\begin{enumerate}

\item Things were good.

\item Then they got bad.

\item But YOU can make them good again.

\end{enumerate}

Not to rock the boat, I too use this story. It generates three big paragraphs.

The first paragraph tells how great it was a long time ago when everyone lived
in harmony.

The second one tells how the world was broken or destroyed or there was some
kind of cataclysmic event that ruined the civilization and let the demons come
into the world, where they had to be destroyed, thereby ruining the world
completely.

Then the third one talks about how the demons are stirring and how they need
to be destroyed by bringing the races together blah blah blah blah...

You get the idea. It's the same crap that everybody else uses, but it has to
be there to set up the world so the players can have some framework to work
with. But it isn't the biggest deal in the world.

The most important thing is that the story is generated after the deities and
ancient races and organizations and societies and the areas in the world, so
it can use that data to personalize the story for this particular
instantiation of the game.

Rememeber, the point isn't to make a particular instantiation of the Heroes'
Journey story, it's to make something that tries to generate all of those
stories (or as many as possible).
