\chapter{Player Characters}


\section{Introduction}

Players are things but they are different from the other things in the
game. The main reason for this is that players have a large amount of
data that only they store, whereas other things fake the data. As an
example, players have actual attributes/stats, and these stats are
accessed with the get\_stat() function, that returns a constant for
nonpcs but a constant for pcs. In a truly object-oriented world, this
would be achieved with a virtual function in the base class Thing that
returns a constant value, and then the value would be calculated for
real in class Player, but this method is close enough.

The thing that distinguishes players in the code is that they have a
pointer to a non NULL pc struct, and that struct is not fake\_pc. The
point behind fake\_pc is to give non-players a piece of data that can
be dereferenced on the off chance that checks aren't made correctly in
the code. Again, in a truly OO world, this would be taken care of
automatically, but this is close enough.

\section{The pc\_data struct}


The player data structure has the following members:

\begin{itemize}
	

\item  PCDATA *next;             This is used to keep a free list of
pc structs for reuse.     

\item  char **curr\_string;            The current string we are
editing...this is a char ** so we can access the address where the
reference to the string is stored to fix that. 

\item  void *editing;                This is what the player is
editing...it could be a THING, a spell, a race, an alignment, or a
society. These are all in the same struct so the player cannot edit
more than one thing at once.

\item  NOTE *note;                This is the note that we are
editing.
 
\item  int curr\_note;            This is the time of the last note we
read. It's not the note number so that the notes can be deleted. 

\item  short stat[STAT\_MAX];         Only players get actual stat
  values. The other things get this faked by get\_stat These range
  from 3 to 32.

\item  short guild[GUILD\_MAX];       These are the tiers a player can
  advance in guilds. There are 10 guilds approx, and 5 tiers and it
  costs N guild points to advance to tier N. The players get 6 guild
  points per remort, so a completed character can have 4 guilds at
  full tier, or all 10 at 3 tiers or some combination in between.

\item short parts[PARTS\_MAX];       These are the numbers of body parts
  that players have. This may be added to all things at some future
  date, but for right now only players have these stats.

 \item  short armor[PARTS\_MAX];      Things have their armor faked too, but
  players get actual armor numbers when they wear eq. The data is
  stored here.

\item  short implants[PARTS\_MAX];    Players can also get implants that
  improve their body parts. In general, head implants make the player
  better at magical things, arm implants make them hit harder, body
  implants add armor, and leg implants let them kick harder and move farther.

\item  int aff[AFF\_MAX];         As usual, affects on things get
  ignored unless they're flags. Affects like kickdam or hitroll have
  no effect on anything other than players. Those modifiers are added
  up and stored here, and they get used to speed up some calculations.

\item  char *email;           This is just the email for the player.

\item  char *pwd;                  This is the encrypted password for
  the player.

\item  char *big\_string;           Big strings get sent to the
players a page at a time using something like a wimpy version of
  more. If the player is in the mode to receive one of these strings,
  this is where the string is stored (and then freed once the player
  is done with it.)

\item  char *big\_string\_loc;        This is the current location of
  the big string this player is viewing.


\item  char *title;                 This is the title that shows up
when a player appears on the who list.


\item  char *prompt;             This is the prompt the player sees walking
  around. It is configurable using the same variables as in scripts
  and score and tset.

\item  int pcflags;             These are some internal flags used
  when players do special things. For example, PCF\_EDIT\_CODE is a
  special flag used if the player is editing code instead of a note or
  description using the little line editor. PCF\_SEARCHING and
  PCF\_TRACKING are used since when a player enters a new room, they
  have a chance to track or search after a delay, and these two flags
  tell whether the player is doing either of these things.

\item  unsigned char race;           This is the race of the character
  taken from the race list.


\item  short remorts;          This is how many remorts this pc
has. Remorting is expensive. You go back to level 1 and lose all of
your skills, but you do gain some stats and some new spells and
skills, and a few extra hps. This builds up over time and is very
useful in the long run even if it sucks short term.


\item  short practices;            Characters must practice skills at
a trainer to learn them up to a point, then they must use the skills
to get them perfect. This is the number of practices a character
has. This is almost never a problem except when a character is really
stupid.


\item  short wait;                  Some things cause lag for
players. The lag gets put here. Things don't get lagged, and admins
can turn off lag by typing config holyspeed to turn on holyspeed.

 \item short ticks;        Some commands like casting and backstabbing
 and investigating corpses have pre-activity lag. This lag gets set
 here. The way this works is that if the player does a skill with
 pre-lag, the ticks get set and they get put into a different position
 to start doing whatever command it is. If they enter a command before
 the ticks are up (or if their concentration is broken) then the ticks
 are set to 0 and they stop trying to do the command. If the ticks get
 down to 0, the players attempt the command.         

 \item short max\_mana;           This is the maximum amount of mana a
 player can store. It is dynamically generated.

 \item short mana;                   This is the current natural mana
 the player has.
 
\item	int pagelen; This is how many lines of text the player can see
 on one screen.

 \item int in\_vnum;                  This is the vnum of the last
 room the player was in. It's used so that if the game reboots while
 the player is in a boat or vehicle, they get put back into a
 ``legit'' room that will exist. It's also used so that if a
 battleground is going on and the player quits or loses link in the bg
 or the game dies, the player ends up in the last room they were in
 before the battleground.

 \item int bank; This is how much money the player has in the bank.

 \item int hostnum;                This is the IP address of the
 player, used to detect cheating.


 \item short cond[COND\_MAX];         These are the conditions a
 player has. Mainly hunger and thirst are used. Drunkenness may be
 added at some point, but it doesn't do much now. When a player eats
 or drinks, his condition in hunger or thirst goes up, and if he
 doesn't eat or drink it goes down until it gets negative and he takes
 regular damage.

\item  unsigned char prac[MAX\_SPELL]; These are the percentages that
 players have learned certain skills. Remember that the max number of
 skills/spells etc... that a player can learn is based on race,
 alignment, and remorts. The percentages go from 0 to 100, and a
 player must have at least 30 pct learned to have it improve through
 use. The max percentage that a player can practice a spell to is 65,
 then it must be learned through use.


 \item short clan\_num[CLAN\_MAX];    There are many different kinds
 of clans: clans, sects, schools. A player can belong to one thing of
 each kind. This is useful if you want several kinds of clans that
 players can join.

\item  THING *storage[MAX\_STORE];      These are items that get
 stored in the bank for the character. It costs a percentage of the
 item value to store and unstore the item, and the most valuable items
 cannot be stored so that people can't keep their items out of pkill.



\item  THING *reply;                When the player tells to another
 thing, this variable gets set to that thing so that the reply command
 works.

\item  TROPHY *trophy[MAX\_TROPHY]; When a player pkills someone, they
 get a trophy of that person. Trophies store the level, align,
 remorts, race and number of times someone is killed. They are sorted
 with remorts taking priority over levels, so there's no way to remort
 and die to get trophies of you ``cleared'' anymore.

\item  char *aliasname[MAX\_ALIAS];    Characters can store aliases. These
  are a list of aliases this player has on his character.

\item  char *aliasexp[MAX\_ALIAS];   These are the alias expansions used by
  the player. Note that you can have variables of the form @1, @2, @3,
  up to @5 in aliases so that you can have more general aliases.

 \item int pk[PK\_MAX]; This stores the pk data for the player. There
 are all kinds of pk data, number of pkills, average level, number of
 times pkilled...and so forth. These items get used to generate the pk
 rankings for the pk command.

 \item int exp;              This is the amount of exp the character
 has. Note that exp to level is from 0 for each level, so that if
 level N says 1 million exp, you really need to get 1 million exp
 after level N-1 to move up. You can advance several levels all at
 once if you manage to kill something buffed at low levels.
        

 \item int fight\_exp;                This is how much exp the player
 has gained from fighting since the last mob kill. It's not a really
 important value...just there to give players an idea of what they're
 getting for fighting a creature.


\item  THING *guarding;              Players using the guard skill on
 other players have this set so that it's a little faster in combat to
 look up who's guarding who.

\item  char *ignore[MAX\_IGNORE];      This is a list of names that
 cannot send channeled messages to this character. It should block ALL
 channels, since they all go through the same function.
 
\item  char temp[STD\_LEN];          When a character is doing a timed
 command, the command has to get stored somewhere, so here is the
 place. When the ticks are up, this command gets executed.


\item  VALUE *qf;           These are a list of questflags (values) that
  the character has to track his progress doing quests. Much nicer
  than having actual flags.

\item  unsigned short off\_def;       This is a number between 0
  (defensive) and 100 (offensive) or in the middle 50 (normal) that
  determines if the character is more focused on offense, defense, or
  a balance.

 \item int no\_quit;                  This is the number of game hours
 (minutes roughly) that a character must wait before he can log
 out. This is to prevent twinks from autofleeing and quitting instantly.

 \item int time;                    This is how many seconds the
 character has played total.

 \item int login;           This is the time the player last logged in.        

 \item int wimpy;             The character will try to autoflee if he
 reaches this many hps. This must be at least 1/3 of the character's
 total hps.

 \item bool channel\_off[MAX\_CHANNEL]; This is a list of channels the
  character is not listening to. These can be toggled by typing
  <channelname> off.

 \item int sback\_pos[MAX\_CHANNEL];  This is a list of scrollback
 positions the character is currently at for channels.

\item  SCROLLBACK *sback[MAX\_CHANNEL][MAX\_SCROLLBACK]; These are recorded
  channel messages the player can see. They are truncated because of
  space and memory concerns over spam allocations. They should
  technically be dynamically allocated buffers.

 \item int pfile\_sent;            This is the time when a copy of the
 character's pfile was sent to the email address provided. It only
 works once every couple of days to try to cut down on annoying people.

 \item int quest\_points;             This is how many quest points
 the character has. This does nothing right now.

\item  bool using\_kde;               This tells whether or not the player
  is using the kde client to log in. If this is set, the player sees
  some extra stuff like a map and some data on the client screen.

 \item int kde\_update\_flags;        These are some flags that
 determine which things get updated on the kde client this tick.

\item  THING *pets;                 These are just used to load pets into
  the world temporarily when players get logged in and before they
  actually get put into a room.

 \item int warmth;             This is the warmth the character has
 from the clothing he is wearing. It doesn't do much now, but it may
 if weather is ever improved.      

\item  MAPGEN *mapgen;            This is a generated map admins can use to
  make new zones faster.  


 \item int gathered[RAW\_MAX]; Players can only mine/gather etc... a
 certain number of raw materials at one time. This saves over reboots,
 so it's a limit on how often the player can gather over time so they
 must spend time online to gather more raw materials.

  bool battling;                This tells if the player will battle
  during the next battleground.


 \item int last\_saved;        This tells the last time the character
 was saved so spam saving isn't possible.     





\end{itemize}















