\chapter{Rewards}

Because the players can affect the world, and because they do things that help
their allies and hurt their enemies, it makes sense to reward them for their
good actions.

There's automatic reward code that gives money, exp and quest points to
players that do good things.

The idea is that a player performs some action that helps or hurts societies,
and it gets recorded. There are several kinds of rewards, so the function call
is society\_add\_reward (THING *th, THING *to, int reward\_type, int amount);

\begin{itemize}

\item th: the player who did the action.

\item to: the thing acted upon. It could be a mob that was healed/killed. It
  could be a room where a player built or razed, and it could be an object
  that the player gave to a society member. This is stored on the player so
  that when the player gets the reward later, this can be referenced (although
  it may be gone by then...which is ok. It just means that the player gets a
  generic reward message.

\item reward\_type: What the player is getting rewarded for. These will be
  explained below.

\item amount: A positive integer from 1-1000 currently. 

\end{itemize}

The player performs a good action and gets the reward recorded, then every
game hour, the amount of the award goes down by 1, so you have to get to a
society member fast to get the full or almost full reward given to you. The
rewards are handed out randomly in the fast\_update() function where you get
one reward type at a time per reward update. It is unlikely that players will
have lots of different rewards pending at once, so this should be
sufficient. I just didn't want to have the rewards being instantaneous.

Any non-child society member can reward a player when they're in the same room.

\section{Reward Types}

There are many kinds of rewards given out for lots of different kinds of
behavior. 

\begin{itemize}

\item play: Playing with children gives you rewards. This doesn't actually
  affect game mechanics for now.

\item build: Building a friendly society city gives you rewards. This is
  useful because cities give combat bonuses to allies.

\item raze: Razing an enemy city. Again, this is useful because cities give
  combat bonuses.

\item heal: Healing/curing friendly society members gives you a reward for
  obvious reasons.

\item kill: Killing enemies gives you rewards for obvious reasons.

\item kidnap: Rescuing a kidnap victim gives you big rewards. In fact, it
  gives rewards to EVERYONE standing in the same room when the reward was
  given out. Could be useful for plevelling newbies.

\item diplomacy: This skill makes societies hate you less and have a larger
  affinity toward your alignment. This makes them more likely to stay or
  switch to your side, and they get cheaper to bribe.

\item inspire: Raises morale for your side which makes them work and fight
  harder.

\item demoralize: Lowers morale of your enemies making them work and fight
  worse.

\item package: Delivering a package given by a leader as a response to "news
  delivery" gives you big rewards because it's a timed mission where enemies
  will attack you, and if you fail, your side has lost a lot of resources over
  nothing.

\end{itemize}
