\chapter{Skills}


\section{Introduction}

Skills is a misleading name for this section. There are really
several kinds of ``skills'' called skills, spells, proficiencies,
traps, and poisons. Each race has a certain number of each type of
skill it can learn, and this number is modified by the alignment the
player has chosen, and the number of remorts that the player
has. Generally, spells are cast, traps are made and set (not done), poisons are
mixed and used on weapons (not done), and skills and proficiencies are
either actively used or passively checked. Skills and proficiencies
are split up into combat skills (skills) and non-combat skills
(proficiencies). 

\section{Mana  and Gems}

To cast a spell, you must have enough mana and you must be able to
cast a spell of that power. Spells have mana of different elemental
types, and if you have enough mana in any one type, you can cast the
spell. Spells also have levels, and the max level you can cast is
based on your stats and guilds and remorts and implants. You can
increase your casting level by using gems. A gem is an item that
stores magical power and boosts the max level of power you can
cast. Just like regular mana, gems require time to recharge, but they
can also increase your casting power dramatically. 

\section{Using skills}

Many skills are linked to commands you must type in to use them. For
example, kick, bash, investigate, sneak. When you attempt to use these
commands, the skill gets checked vs. your ability to see if you
succeed. Mobs have automatic chances to use skills based on their
levels. If you want to change these automatic chances, look at the
function called check\_spell() to tweak this.


